Production study: Methodological details

### 2021-01-19 docs/ExperimentalDetails/Experimental_details.tex
:memo: - edit ExperimentalDetails

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Methodological Details - Blue Bananas},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{expex}
\lingset{everygla=,everyglpreamble=\it}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Methodological Details - Blue Bananas}
\author{}
\date{\vspace{-2.5em}9/17/2020}

@@

\hypertarget{research-questions}{%
\section{Research Questions}\label{research-questions}}

This study investigates whether speakers modify the speech signal to
signal atypical referents? Concretely, we ask whether the suprasegmental
profile of an utterance with an atypical referent like ``blue banana''
is different from one with a typical referent such as ``yellow banana''.

@@

In this production study, participants interacted with the experimenter.
Participants had to verbally instruct the experimenter to select a
specified target object out of four visually presented objects. The
non-target objects differed from the target with respect to their
colour, their identity, or both. Objects were referred to using noun
phrases that consist of a modifier denoting colour and a modified object
(e.g.~yellow banana, blue banana). These adjective-noun combinations
differed with respect to the typicality of their combination. The
adjective-noun combinations were either typical, medium typical, or
atypical (as established in a norming study, see
\protect\hyperlink{adds}{additional materials}). Some combinations were
typical such as yellow banana, red tomato, or yellow apricot; some were
atypical such as red banana, purple tomato, or blue apricot; and some
were somewhat in the middle like brown banana, green tomato, and red
apricot.

@@

Participants were seated in front of a computer screen. The experimenter
sat at the opposite side of the table in front of another computer
screen. The participants and the experimenter could see neither each
other nor each others' screens. The experiment consisted of two phases:
a familiarization phase and a test phase.

In a familiarization phase, participants saw one object per trial. In
order to advance to the next trial, participants had to read out loud
the corresponding noun phrase (e.g.~blue banana). During this phase,
participants had to name all atypical colour-object targets that we used
in the test phase of the experiment alongside their typical
counterparts. For example, if red banana (atypical) was an experimental
target, participants were presented with both the red banana (atypical)
and the yellow banana (typical). This familiarization phase was included
in order to ensure that participants can relate typical and atypical
colour-object combinations to each other.

After the familiarization phase, participants entered the test phase. On
each trial in the test phase, participants first saw four coloured
objects in the top left, top right, bottom left, and bottom right of the
screen, respectively. One of the object served as the target; another
served as the competitor; and the remaining two served as unrelated
distractors. The position of the visual stimuli was randomized for each
trial and each participant. In the center of the screen, a black cube
was displayed which could be moved by the experimenter. The participants
were asked to instruct the experimenter to move the cube onto one of the
four pictures.

Each test trial consisted of two parts. The `trigger' instruction and
the `test' instruction. After the preview of all images was displayed
for 1500 ms, the competitor object was visually highlighted by an arrow,
and a `trigger' instruction was orthographically presented below the
cube (see Figure \#):

@@

{\centering \includegraphics[width=0.6\linewidth]{sample_experiment_screen_competitor}

\caption{Example screen at the beginning of a trials. The discourse setting trigger sentence is displayed in the middle below the cube. The participants have to instruct the experimenter to move the cube on top of the green sunglasses (indicated by an arrow).}\label{fig:image1}
\end{figure}

The trigger instruction was supposed to create a discourse context, such
that both the color and the competitor object were introduced as
background information into the discourse (here: green and sunglasses).
The trial proceeded when the experimenter has moved the cube onto the
respective referent. The sentence and the arrow disappeared.
Subsequently, the target object was visually highlighted by an arrow and
the test instruction was presented containing the target object (see
Figure \#):

@@

{\centering \includegraphics[width=0.6\linewidth]{sample_experiment_screen_target}

@@

\caption{Example screen after the trigger sentence. The target sentence is displayed in the middle of the screen. Now, participants have to instruct the experimenter to move the cube on top of the green avocado (indicated by an arrow).}\label{fig:image2}
\end{figure}

@@

Using the trigger instruction to set up the discourse context enabled us
to manipulate the focus structure of the target sentence. If the color
of the competitor and the target are of the same object type but
differed in color (e.g.~yellow banana --\textgreater{} blue banana), the
color adjective was discourse-pragmatically focused (henceforth the
Adjective Focus (AF) condition). If the objects differed but not their
color (e.g.~yellow banana --\textgreater{} yellow tomato), the noun was
in focus (henceforth the Noun Focus (NF) condition). If both the color
and the object differed (e.g.~yellow banana --\textgreater{} blue
tomato), the whole noun phrase was in focus (henceforth the Double Focus
(DF) condition).

We are interested in the Noun Focus condition, only. AF and DF served as
filler trials. There were 15 NF trials, 10 AF and 10 DF trials each.
Each trial occurred two times per participants, yielding a total of 70
trials per participant.

The experimental program (implemented as a browser-based application) is
available here: \url{https://github.com/SBRitter/fruits-production}.

@@

\hypertarget{adds}{%
\section{Additional Information}\label{adds}}

Here we specify in detail, how stimuli were selected for the above
described production study.

@@

The selection of stimuli was informed by a norming study for which we
selected twenty fruits or vegetables (henceforth: FOODs) and four other
referents (henceforth: NON-FOODs). Selection criteria were mainly
informed by visual discriminability and phonemic composition, such that
nouns denoting our objects were supposed to contain few voiceless
segments.

For each of the FOODs and NON-FOODs, we created colored versions by
image manipulation using \href{http://www.gimp.org}{gimp}. Images of all
individual objects were collected from internet databases containing
copyright-free high-resolution images
(e.g.~\href{https://pixabay.com}{Pixabay}). During the manipulation
process, the background of the pictures was replaced by a white
background. In order to change the color of the objects, a layer in the
respective color was created and overlaid on top of the original image
using ``Color'' as the layer mode. The hexadecimal color codes for the
colors that we used are listed below:

\noindent Blue: \#29429f\\
\noindent Green: \#3f8535\\
\noindent Red: \#9d1c1c\\
\noindent Purple: \#6d1b79\\
\noindent Brown: \#2a1d11\\
\noindent Yellow: \#e3c917\\
\noindent Orange: \#ff8400\\
\noindent Orange (potatoes): \#ff6600\\
\noindent Black: \#000000\\
\noindent Grey: \#FFFFFF

@@

The colors for the twenty FOODs were subjectively selected to reflect a
range of compatibilities with the respective objects, e.g.~a typical
banana is yellow; a less typical banana could be brown (too ripe) or
green (not yet ripe); and an atypical banana could be blue. Every FOOD
was presented in four of nine manipulated colors alongside its original
image. The four NON-FOODs were subjectively selected to represent color
agnostic objects which can and do come in a variety of colors. All
objects were chosen such that their respective German nouns were either
feminine singular (e.g.~``die Banana'' `the banana') or plural
(e.g.~``die Trauben'' `the grapes').

@@

\hypertarget{the-norming-study}{%
\subsection{The norming study}\label{the-norming-study}}

One-hundred German native speakers participated in a norming study using
the crowd sourcing platform
{[}Prolific{]}{]}(\url{https://www.prolific.ac}). We presented all
objects in different colors to our participants. FOODs came in 5 colors
per object, NON-FOODs came in all nine colors. Participants were
instructed to rate how typical they think the color for each object was,
using a smooth slider ranging from 1 to 100 (see javascript for the
experimental design here:
\url{https://github.com/stelaseldano/colour-typicality-norming}. The
results of the norming study can be retrieved here:
\url{https://osf.io/znpg5/}.

\hypertarget{selection-of-experimental-stimuli}{%
\subsection{Selection of experimental
stimuli}\label{selection-of-experimental-stimuli}}

The norming ratings were subsequently used to select appropriate stimuli
for the production study according to the following procedure:

@@

For the critical focus condition (NF) this selection procedure resulted
in 15 target FOODs (5 colors x 3 typicality categories):

\noindent atypical:\\
\noindent Yellow cherry (Gelbe Kirsche)\\
\noindent Green carrot (Grüne Möhre)\\
\noindent Red cucumber (Rote Gurke)\\
\noindent Orange grapes (Oangene Trauben)\\
\noindent Brown pepper (Braune Paprika)

\noindent medium:\\
\noindent Yellow peas (Gelbe Erbsen)\\
\noindent Green tomato (Grüne Tomate)\\
\noindent Red apricot (Rote Aprikose)\\
\noindent Orange potatoes (Orangene Kartoffeln)\\
\noindent Brown banana (Braune Banane)

\noindent typical:\\
\noindent Yellow lemon (Gelbe Zitrone)\\
\noindent Green green beans (Grüne Bohnen)\\
\noindent Red strawberry (Rote Erdbeere)\\
\noindent Orange mandarine (Orangene Mandarine)\\
\noindent Brown walnut (Braune Walnuss)

@@

\noindent Yellow sunglasses (Gelbe Sonnenbrille)\\
\noindent Yellow socks (Gelbe Socken)\\
\noindent Yellow clothes peg (Gelbe Wäscheklammer)\\
\noindent Green sunglasses (Grüne Sonnenbrille)\\
\noindent Green socks (Grüne Socken)\\
\noindent Green Paper clip (Grüne Büroklammer)\\
\noindent Red socks (Rote Socken)\\
\noindent Red paper clip (Rote Büroklammer)\\
\noindent Red clothes peg (Rote Wäscheklammer)\\
\noindent Orange socks (Orangene Socken)\\
\noindent Orange paper clip (Orangene Büroklammer)\\
\noindent Orange clothes peg (Orangene Wäscheklammer)\\
\noindent Brown sunglasses (Braune Sonnenbrille)\\
\noindent Brown paper clip (Braune Büroklammer)\\
\noindent Brown clothes peg (Braune Wäscheklammer)

The pairing of targets and competitors was randomized for each subject,
with the only constraint that the competitor had the same color as the
target object (e.g.~if blue banana was the target, blue grapes is a
possible competitor) (see randomization section below for more details).

For the filler focus conditions (AF + DF), we distributed FOODs and
NON-FOODs to fulfil the following constraints: (a) each target FOOD
appeared 3 times and (b) each color appeared as a traget color 14 times.

For all three focus conditions, a subset of distractors was selected
from the initial superset: Distractors were either FOODS that were
neither used as targets nor as competitors (avocado, egg plant, pear,
zucchini) or NON-FOODs. Within each given trial, the distractors did
neither share color nor object identity with neither the target nor the
competitor.

Overall, the sets of target, competitors and distractors are combined
such that all five colors occurred equally often throughout the
experiment (= 28 times).

\hypertarget{randomization}{%
\section{Randomization}\label{randomization}}

For each trial, targets, competitors, and distractors are chosen in a
randomized way, fulfilling the following requirements:

\begin{itemize}
\item
  NF-Condition:\\
  Color of Target = Color of Competitor\\
  Object of Target != Object of Competitor
\item
  AF-Condition:\\
  Color of Target != Color of Competitor\\
  Object of Target = Object of Competitor
\item
  DF-Condition:\\
  Color of Target != Color of Competitor\\
  Object of Target != Object of Competitor
\item
  All conditions:\\
  Both distractors have to differ in color and object identity from both
  target and competitor.
\end{itemize}

The competitor and distractors are matched iteratively to a target. One
competitor could only be used twice or a third time if all other
competitors that were left had been considered and prove inappropriate.

The set of four objects (quadruple) was formed separately for each
condition. After the three lists had been compiled, they were merged.
The sequence of quadruples was randomized such that the following
requirements were met:

The python code for the randomization is available here:
\url{https://github.com/SBRitter/fruits-randomization}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Production study: Methodological details},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{expex}
\lingset{everygla=,everyglpreamble=\it}
\usepackage[export]{adjustbox}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Production study: Methodological details}
\author{}
\date{\vspace{-2.5em}9/17/2020}

@@

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\hypertarget{overview}{%
\section{Overview}\label{overview}}

This document describes the methodological details of the production study, which generated the data your team will analyze for the \emph{Many Speech Analyses} project.

Although the details of the norming study (run prior to the production study, see OSF Wiki) might not be directly relevant for your analysis, we provide an outline of the norming study methods in Appendix \ref{norming} for context.

\hypertarget{research-question}{%
\section{Research question}\label{research-question}}

This study investigated whether speakers modify speech to signal atypical referents.
Concretely, we asked whether the acoustic profile of an utterance with an atypical referent like \emph{blue banana} is different from one with a typical referent such as \emph{yellow banana}.

@@

In this production study, participants interacted with the experimenter in German.
Participants were seated in front of a computer screen.
The experimenter sat at the opposite side of the table in front of another computer screen.
The participants and the experimenter could not see each other nor each others' screens.
The participants had to verbally instruct the experimenter to select a specified target object out of four visually presented objects.
The non-target objects differed from the target with respect to their color, their identity, or both.
Objects were referred to using noun phrases that consist of a modifier denoting color and a modified object (e.g.~\emph{Gelbe Zitrone} `yellow lemon', \emph{Rote Gurke} `red cucumber', \emph{Rote Socken} `red socks').
These color-object combinations differed with respect to their typicality.
The color-object combinations were either typical (e.g.~\emph{Orangene Mandarine} `orange mandarin'), medium typical (e.g.~\emph{Grüne Tomate} `green tomato'), or atypical (e.g.~\emph{Gelbe Kirsche} `yellow cherry').
The typicality rating of each color-object combination was established, prior to the production experiment, with a norming study (see Appendix \ref{norming} for details).
Figure \ref{fig:image0} illustrates combination typicality for a banana, a tomato, and an apricot.

@@

The experiment consisted of two phases: a familiarization phase and a test phase.

In a familiarization phase, participants saw the picture of one object per trial.
In order to advance to the next trial, participants had to read out loud the corresponding color-object combination.
During this phase, participants had to name all atypical color-object targets that we used in the test phase of the experiment alongside their typical counterparts.
For example, one of the experimental targets was a red banana (atypical).
The participants were presented with the picture of both a red banana (atypical) and a yellow banana (typical).
This familiarization phase was included in order to ensure that participants could relate typical and atypical color-object combinations to each other.

After the familiarization phase, the participants entered the test phase.
On each trial in the test phase, the subjects first saw four colored objects in the top left, top right, bottom left, and bottom right of the screen, respectively.
One of the object served as the target; another served as the competitor; and the remaining two served as unrelated distractors.
The position of the visual stimuli was randomized for each trial and each participant (see Appendix \ref{random}).
In the center of the screen, a black cube was displayed, which could be moved by the experimenter.
The participants were asked to instruct the experimenter to move the cube on top of one of the four depicted object.

{\centering \includegraphics[width=0.6\linewidth,frame]{sample_experiment_screen_competitor}

\caption{Test trial: Example screen of a trigger instruction. The trigger instruction is displayed in the middle below the cube. Here, the participants have to instruct the experimenter to move the cube on top of the green sunglasses (indicated by an arrow).}\label{fig:image1}
\end{figure}

Each test trial consisted of two parts.
The trigger instruction and the test instruction.
After the preview of all images was displayed for 1500 ms, the competitor object was visually highlighted by an arrow, and a trigger instruction was presented in writing below the cube (see Figure \ref{fig:image1}).
The trigger instruction was constituted by the frame sentence \emph{Du sollst den Würfel auf X ablegen}.
\emph{X} was one of the adjective-noun combinations.

For example, the subjects saw the following sentence:

@@

The trigger instruction was supposed to create a discourse context, such that both the color and the competitor object were introduced as background information into the discourse (in this example: \emph{green} and \emph{sunglasses}).
The trial proceeded when the experimenter has moved the cube onto the respective referent, at which point the sentence and the arrow would disappear.

Subsequently, the target object was visually highlighted by an arrow and the test instruction containing the target object was presented (see Figure \ref{fig:image2}).
This time the frame sentence was \emph{Und jetzt sollst du den Würfel auf X ablegen}, where \emph{X} was again an adjective-noun combination.
To continue the example above, the subject would see the following test instruction:

@@

{\centering \includegraphics[width=0.6\linewidth,frame]{sample_experiment_screen_target}

@@

\caption{Test trial: Example screen of the test instruction. The target sentence is displayed in the middle of the screen. Now, participants have to instruct the experimenter to move the cube on top of the green avocado (indicated by an arrow).}\label{fig:image2}
\end{figure}

@@

Using the trigger instruction to set up the discourse context enabled us to manipulate the focus structure of the target sentence.
If the color of the competitor and the target are of the same object type but differed in color (e.g.~\emph{yellow banana} vs \emph{blue banana}), the color adjective was discourse-pragmatically focused (henceforth the Adjective Focus (AF) condition).
If the objects differed but their color did not (e.g.~\emph{yellow banana} vs \emph{yellow tomato}), the noun was in focus (henceforth the Noun Focus (NF) condition).
If both the color and the object differed (e.g.~\emph{yellow banana} vs \emph{blue tomato}), the whole noun phrase was in focus (henceforth the Adjective/Noun Focus (ANF) condition).

There were 15 NF trials, 10 AF and 10 ANF trials each.
Each trial occurred two times per participants, yielding a total of 70 trials per participant.

The program (implemented as a browser-based application) used for the experiment is available at the following link: \url{https://github.com/SBRitter/fruits-production}.

@@

\hypertarget{random}{%
\section{Randomization}\label{random}}

In the production study, the combination of target/competitor/distractors were randomized in each trial.
Randomization was restricted to satisfy the following criteria, depending on the condition.

\begin{itemize}
\tightlist
\item
  NF-Condition:

  \begin{itemize}
  \tightlist
  \item
    Color of Target = Color of Competitor.
  \item
    Object of Target != Object of Competitor.
  \end{itemize}
\item
  AF-Condition:

  \begin{itemize}
  \tightlist
  \item
    Color of Target != Color of Competitor.
  \item
    Object of Target = Object of Competitor.
  \end{itemize}
\item
  DF-Condition:

  \begin{itemize}
  \tightlist
  \item
    Color of Target != Color of Competitor.
  \item
    Object of Target != Object of Competitor.
  \end{itemize}
\item
  Distractors, all conditions:

  \begin{itemize}
  \tightlist
  \item
    Both distractors have to differ in color and object identity from both target and competitor.
  \end{itemize}
\end{itemize}

The competitor and distractors were matched iteratively to a target.
One competitor could only be used twice or a third time if all other competitors that were left had been considered and proved inappropriate.

The set of four objects (quadruple) was formed separately for each of the three focus conditions.
After the three lists had been compiled, they were merged.
The sequence of quadruples was randomized such that the following requirements were met:

The python code for the randomization is available at the following link: \url{https://github.com/SBRitter/fruits-randomization}.

\hypertarget{norming}{%
\section{Norming study}\label{norming}}

Here we specify in detail the selection of the stimuli for the production study.

@@

The selection of stimuli was informed by a norming study for which we have chosen twenty fruits or vegetables (henceforth: FOODs) and four other referents (henceforth: NON-FOODs).
Selection criteria were mainly informed by visual discriminability and phonemic composition, such that nouns denoting our objects were supposed to contain as few voiceless segments as possible.

For each of the FOODs and NON-FOODs, we created colored versions by image manipulation using \href{http://www.gimp.org}{Gimp}.
Images of all individual objects were collected from internet databases containing copyright-free high-resolution images (e.g.~\href{https://pixabay.com}{Pixabay}).
During the manipulation process, the background of the pictures was replaced with a white background.
In order to change the color of the objects, a layer in the respective color was created and overlaid on top of the original image using \texttt{Color} as the layer mode.
The hexadecimal color codes for the colors that we used are listed below:

\begin{itemize}
\tightlist
\item
  Blue: \#29429f
\item
  Green: \#3f8535
\item
  Red: \#9d1c1c
\item
  Purple: \#6d1b79
\item
  Brown: \#2a1d11
\item
  Yellow: \#e3c917
\item
  Orange: \#ff8400
\item
  Orange (potatoes): \#ff6600
\item
  Black: \#000
\item
  Grey: \#fff
\end{itemize}

@@

The colors for the twenty FOODs were subjectively selected to reflect a range of compatibilities with the respective objects, e.g.~a typical banana is yellow; a less typical banana could be brown (too ripe) or green (not yet ripe); and an atypical banana could be blue.
Every FOOD was presented in four of nine manipulated colors alongside its original image.
The four NON-FOODs were subjectively selected to represent color agnostic objects which can and do come in a variety of colors.
All objects were chosen such that their respective German nouns were either feminine singular (e.g.~\emph{die Banana} `the banana') or plural (e.g.~\emph{die Trauben} `the grapes').

@@

\hypertarget{participants-and-procedures}{%
\subsection{Participants and procedures}\label{participants-and-procedures}}

A hundred German native speakers participated in a norming study using the crowd sourcing platform \href{https://www.prolific.ac}{Prolific}.
We presented all objects in different colors to our participants.
FOODs came in 5 colors per object, NON-FOODs came in all nine colors.
Participants were instructed to rate how typical they thought the color for each object was, using a smooth slider ranging from 1 to 100 (see javascript for the experimental design here: \url{https://github.com/stelaseldano/colour-typicality-norming}).
The results of the norming study can be retrieved here: \url{https://osf.io/znpg5/}.

\hypertarget{selection-of-experimental-stimuli-for-the-production-study}{%
\subsection{Selection of experimental stimuli for the production study}\label{selection-of-experimental-stimuli-for-the-production-study}}

The typicality ratings were subsequently used to select appropriate stimuli for the production study according to the following procedure:

@@

For the noun focus condition (NF) this selection procedure resulted in 15 target FOODs (5 colors x 3 typicality categories):

\begin{itemize}
\tightlist
\item
  Atypical:

  \begin{itemize}
  \tightlist
  \item
    Yellow cherry (\emph{Gelbe Kirsche}).
  \item
    Green carrot (\emph{Grüne Möhre}).
  \item
    Red cucumber (\emph{Rote Gurke}).
  \item
    Orange grapes (\emph{Oangene Trauben}).
  \item
    Brown pepper (\emph{Braune Paprika}).
  \end{itemize}
\item
  Medium:

  \begin{itemize}
  \tightlist
  \item
    Yellow peas (\emph{Gelbe Erbsen}).
  \item
    Green tomato (\emph{Grüne Tomate}).
  \item
    Red apricot (\emph{Rote Aprikose}).
  \item
    Orange potatoes (\emph{Orangene Kartoffeln}).
  \item
    Brown banana (\emph{Braune Banane}).
  \end{itemize}
\item
  Typical:

  \begin{itemize}
  \tightlist
  \item
    Yellow lemon (\emph{Gelbe Zitrone}).
  \item
    Green green beans (\emph{Grüne Bohnen}).
  \item
    Red strawberry (\emph{Rote Erdbeere}).
  \item
    Orange mandarine (\emph{Orangene Mandarine}).
  \item
    Brown walnut (\emph{Braune Walnuss}).
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Yellow sunglasses (\emph{Gelbe Sonnenbrille}).
\item
  Yellow socks (\emph{Gelbe Socken}).
\item
  Yellow clothes peg (\emph{Gelbe Wäscheklammer}).
\item
  Green sunglasses (\emph{Grüne Sonnenbrille}).
\item
  Green socks (\emph{Grüne Socken}).
\item
  Green Paper clip (\emph{Grüne Büroklammer}).
\item
  Red socks (\emph{Rote Socken}).
\item
  Red paper clip (\emph{Rote Büroklammer}).
\item
  Red clothes peg (\emph{Rote Wäscheklammer}).
\item
  Orange socks (\emph{Orangene Socken}).
\item
  Orange paper clip (\emph{Orangene Büroklammer})
\item
  Orange clothes peg (\emph{Orangene Wäscheklammer}).
\item
  Brown sunglasses (\emph{Braune Sonnenbrille}).
\item
  Brown paper clip (\emph{Braune Büroklammer}).
\item
  Brown clothes peg (\emph{Braune Wäscheklammer}).
\end{itemize}

For the Adjective and Adjective/Noun conditions (AF + ANF), we distributed FOODs and NON-FOODs to fulfill the following constraints: (a) each target FOOD appeared 3 times and (b) each color appeared as a target color 14 times.

For all three focus conditions, a subset of distractors was selected from the initial superset.
The distractors were either FOODS that were neither used as targets nor as competitors (avocado, egg plant, pear, zucchini) or NON-FOODs.
Within each given trial, the distractors did not share color nor object identity with neither the target nor the competitor.

Overall, the sets of target, competitors and distractors were combined such that all five colors occurred equally often throughout the experiment (i.e.~28 times).
